---
title: "Capstone_Project_bikecase"
author: "Posatron"
output: html_document
date: '2022-06-04'
---

## Call the library you need for this job.
```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(skimr)
library(readr)
library(ggplot2)
```

## Import data file and rename them with simple name.
```{r Import and rename}
trips_2008 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202008-divvy-tripdata/202008-divvy-tripdata.csv")
trips_2009 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202009-divvy-tripdata/202009-divvy-tripdata.csv")
trips_2010 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202010-divvy-tripdata/202010-divvy-tripdata.csv")
trips_2011 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202011-divvy-tripdata/202011-divvy-tripdata.csv")
trips_2012 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202012-divvy-tripdata/202012-divvy-tripdata.csv")
trips_2101 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202101-divvy-tripdata/202101-divvy-tripdata.csv")
trips_2102 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202102-divvy-tripdata/202102-divvy-tripdata.csv")
trips_2103 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202103-divvy-tripdata/202103-divvy-tripdata.csv")
trips_2104 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202104-divvy-tripdata/202104-divvy-tripdata.csv")
trips_2105 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202105-divvy-tripdata/202105-divvy-tripdata.csv")
trips_2106 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202106-divvy-tripdata/202106-divvy-tripdata.csv")
trips_2107 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202107-divvy-tripdata/202107-divvy-tripdata.csv")
trips_2108 <- read_csv("E:/Data Analysts/Case Study/Project_Bike_Case/202108-divvy-tripdata/202108-divvy-tripdata.csv")

```

## Check the column names.
Before merging individual files into one integrated file, Look at the elements of the individual table. This is essential part of the data cleaning process.

```{r Check the colomn names}
colnames(trips_2008)
```
## Take a look at the data table

```{r Look at the data table}
View(trips_2008)
```

## Check the column names and the data format of each column.
```{r Column names and the data Format}
str(trips_2008)
```

## Compare Different column 1.

As you see from the result below, We can see that column end_station_id and column start_station_id have a different data format in several column. From August to November of 2020, it is integer, and after December the format is change to character.
Integration is possible only when those are changed to the same data format.
```{r Compare_df_cols}
compare_df_cols(trips_2008, trips_2009, trips_2010, trips_2011, trips_2012, trips_2101, trips_2102, trips_2103, trips_2104, trips_2105, trips_2106, trips_2107, trips_2108)

```
## Change the integer data format to character.
The target columns are the end station id and start station id columns from August to November.
```{r change the data format}
trips_2008 <- mutate(trips_2008, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips_2009 <- mutate(trips_2009, end_station_id = as.character(end_station_id),
start_station_id = as.character(start_station_id))
trips_2010 <- mutate(trips_2010, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips_2011 <- mutate(trips_2011, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
```

## Compare Different column 2.
Use return = "mismatch" to see if result 0_rows that means there is no mismatch data format anymore.
```{r}
compare_df_cols(trips_2008, trips_2009, trips_2010, trips_2011, trips_2012, trips_2101, trips_2102, trips_2103, trips_2104, trips_2105, trips_2106, trips_2107, trips_2108, return = "mismatch")
```
## Now combine the tables into one integrate file.
Lets integrate several tables into one integrate file.
```{r combine into one file.}
trips_2008_2108 <- bind_rows(trips_2008, trips_2009, trips_2010, trips_2011, trips_2012, trips_2101, trips_2102, trips_2103, trips_2104, trips_2105, trips_2106, trips_2107, trips_2108)
```
## Let check the integrated file.
```{r}
View(trips_2008_2108)
```
## Let check data format again before go to analyst phase.
Becareful data and time if its the POSIXct or POSIXlt format, it is done properly 
```{r}
str(trips_2008_2108)
```

## Check the total number of row and columns.
The first is number of row and latter is the number of columns.
```{r}
dim(trips_2008_2108)
```
## Touch the outline with the summary function.
```{r}
summary(trips_2008_2108)
```
## Create new columns such as month, date, and day of the week for analysis.
```{r Create columns as month, day, and the day of the week}
trips_2008_2108$date <- as.Date(trips_2008_2108$started_at)
trips_2008_2108$month <- format(as.Date(trips_2008_2108$date), "%m")
trips_2008_2108$day <- format(as.Date(trips_2008_2108$date), "%d")
trips_2008_2108$year <- format(as.Date(trips_2008_2108$date), "%Y")
trips_2008_2108$day_of_week <- format(as.Date(trips_2008_2108$date), "%A")

```
## Find the lenght of each trip.
```{r}
trips_2008_2108$ride_length <- difftime(trips_2008_2108$ended_at, trips_2008_2108$started_at)
```
## Check the current work status.
```{r pressure, ECHO=FALSE}
View(trips_2008_2108)
str(trips_2008_2108)

```
## Lets change the data format of ride_lenght to "numeric"

```{r}
trips_2008_2108$ride_length <- as.numeric(as.character(trips_2008_2108$ride_length))
```
## Remove unnecessary values.
For example, values whose ride_length is less than zero. 
And attach a version number to distinguish it. Add version 1.
```{r Remove unnecessary values}
trips_2008_2108_V1 <- trips_2008_2108[!(trips_2008_2108$ride_length<0),]

```
## Final inspection.
```{r pressure, echo=FALSE}
summary(trips_2008_2108_V1)
summary(trips_2008_2108_V1$ride_length)

```
## File Extraction 
Lets take this extracted file, you can analyze in the various way to share your project in Tableau
```{r}
write.csv(trips_2008_2108_V1, "trips_2008_2108.csv")
```

